Parallelization of Push-based System for Molecular Simulation Data Analysis with GPU

### 2016-02-09 paper/paperSource.tex
automatic triggered changes

\subsection{Query modules}
As mentioned earlier, there are two types of functions/queries used
for analysis of MS systems: algebraic or one-body, and holistic or
two-body queries (these in general can be multi-body, but in this
paper we only deal with a two-body functions).

\subsubsection{One-body queries} Most of the query modules in
Table~\ref{tb:queries} (except the SDH and RDF) are not that
involved, only containing computations of fairly simple, one-body
functions. These queries were coded as separate modules in our
system. Each of these modules take few attributes as input (e.g.,
atom selection, frames selection(for the autocorrelation functions),
number of atoms, etc.). The system pushes the data as it becomes
available onto these modules. The queries are being executed on the
selection and are put in a ``ready'' mode, awaiting the next frame's
data. First, the more basic queries, like total mass, are being
computed. The results of such queries are temporary stored (in main
memory) and are available for use anytime a more complex query needs
them.

>>>>

\subsection{Network of queries organization}

Since we assume that our framework should be easily used by a scientist doing MS analysis, it is important to keep queries module as a separate piece of code to make it easily extendable. In order to understand how queries module is organized and the reason for it, we need to understand how CUDA and GPU work.

\emph{\textbf{GPU execution}}. From hardware standpoint, it is important to understand that every GPU device consists of SM (streaming multiprocessors), and each SM has thousands of registers, which can be partitioned among execution threads. Having this in mind, since this is a pure parallel processing power, the goal is to increase hardware throughput and keep highest utilization to exploit every register given. CUDA is an amazing tool that tries to help developers map this hardware capabilities with the actual application layer. In CUDA we have grids, blocks, and threads, where grid is 3 dimensional unit where each cell is a block, and the same relationship applies between blocks and threads\footnote{This explanation doesn't include memory setup in GPU devices, but it is covered in further sections}. Originally, this mapping was perfect for solving image processing problems or other graphics computation, since dimensions and such division (e.g. by pixels to threads) is very inherent. As a language feature, in CUDA we have to define such functions called 'kernel' function. Essentially, it is the function where we are supposed to map every execution thread with data and perform the actual execution, this approach is widely known as SIMD\footnote{SIMD - Single Instruction Multiple Data}. As a convenience we have separated one-body functions with multiple-body functions into separate kernels.

\emph{\textbf{One-body queries kernel}}.
Most of the query modules in Table~\ref{tb:queries} (except the SDH and RDF) are not that involved, only containing computations of fairly simple, one-body functions. These queries were coded as separate modules in our
system. Each of these modules take few attributes as input (e.g.,
atom selection, frames selection(for the autocorrelation functions),
number of atoms, etc.). The system pushes the data as it becomes
available onto these modules. The queries are being executed on the
selection and are put in a ``ready'' mode, awaiting the next frame's
data. First, the more basic queries, like total mass, are being
computed. The results of such queries are temporary stored (in main
memory) and are available for use anytime a more complex query needs
them.
